---
title: "Class 9 Mini Learning Project"
author: "Reina Bassil"
date: "10/30/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
wisc.df <- read.csv("WisconsinCancer.csv", header = TRUE)
```

```{r}
head(wisc.df)
```

Q1.. What type of object is returned from the read.csv() function?
```{r}
class(wisc.df)
```
Data frame

Q2. How many observations (i.e. patients) are in this dataset?
Here we examine data for `r nrow(wisc.df)` patient samples.
```{r}
nrow(wisc.df)
```
569 patients


Q3. How many of the observations have a malignant diagnosis? 
```{r}
cancerous <- table(wisc.df$diagnosis)
cancerous
```
In this data set we have `r cancerous[1]` non cancer and `r cancerous[2]` cancer diagnoses.

In this data set we have `r cancerous["B"]` non cancer and `r cancerous["M"]` cancer diagnoses.



Q4. How many variables/features in the data are suffixed with _mean? 

```{r}
colnames(wisc.df)
#Could count by hand, or use grep
```

```{r}
grep("_mean", colnames(wisc.df))

y <- length(grep("_mean", colnames(wisc.df)))
```

There are `y` variables in the data suffixed by _mean.


Q5. Why do you think we are using the indices 3:32 here?
We want to keep the patient IDs and whether the tumors are benign or malignant out of the data analysis. In addition, there is a final row X that has NA data that could cause errors.

```{r}
wisc.data <- as.matrix( wisc.df[,3:32])
```


The next step in our analysis is to perform principal component analysis (PCA) on wisc.data.

Do we need to scale the data?

```{r}
round( apply (wisc.data, 2, sd), 3)
```


Looks like we need to use `scale=TRUE` here as our data standard deviations are very different. If we run unscaled, the area would drive the whole calculation. Since we are measuring different things with different columns, we need to scale to make comparisons more fair. 

```{r}
#Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp( wisc.data, scale= TRUE)
```



```{r}
summary(wisc.pr)
```

```{r}
x <- summary(wisc.pr)
x$importance

x$importance[,"PC1"]
```

Q7. From your results, what proportion of the original variance is captured by the first principal components (PC1)? 
The first PC captures `r x$importance[2,"PC1"]*100`% of the original variance in the dataset.
```{r}
x$importance[2,"PC1"]*100
```


Q8. How many principal components (PCs) are required to describe at least 70% of the original variance in the data? 
`r which(x$importance[3,] > 0.7)[1]` PCs are required to describe at least 70% of the original variance in the data.
```{r}
which(x$importance[3,] > 0.7)[1]
```

Q9. How many principal components (PCs) are required to describe at least 90% of the original variance in the data? 
`r which(x$importance[3,] > 0.9)[1]` PCs are required to describe at least 70% of the original variance in the data.
```{r}
which(x$importance[3,] > 0.9)[1]
```



Plot PC1 vs PC2 and color by malignant/benign
```{r}
biplot(wisc.pr)
```


Scree Plot
```{r}
plot(wisc.pr)
```

Make diagnosis vector
```{r}
diagnosis <- wisc.df[,2]
```



```{r}
plot(wisc.pr$x[,1:2], col = wisc.df$diagnosis)
#include [,1] and [,2] for PC1 and PC2. It defaults but if you do other PC then use this. 
plot(wisc.pr$x)
```





##Hierarchical clustering


```{r}
plot(hclust( dist(wisc.data)))
```




Use input from our principal component analysis
```{r}
plot(hclust( dist(wisc.pr$x[,1:7]), "ward.D2"))
wisc.pr.hclust <- hclust( dist(wisc.pr$x[,1:7]), "ward.D2")
```

```{r}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
table(grps, diagnosis)
```

```{r}
plot(wisc.pr$x[,1:2], col=grps)

```





```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

