---
title: "Class 11: Structural Bioinformatics 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Download a CSV file from the PDB site
https://www.rcsb.org/stats/summary

```{r}
PDBdata_dist <- read.csv("Data Export Summary.csv")
```


Q. Determine the percentage of structures solved by X-Ray and Electron Microscopy
```{r}
XRay <- (PDBdata_dist$Total[1])/sum(PDBdata_dist$Total)*100
XRay

EM <- (PDBdata_dist$Total[3])/sum(PDBdata_dist$Total)*100
EM

#Alternatively,
ans <- PDBdata_dist$Total/sum(PDBdata_dist$Total) * 100
names(ans) <- PDBdata_dist$Experimental.Method
round(ans,2)
```

Q. What proportion of structures are protein? 
```{r}
round( sum(PDBdata_dist$Proteins)/ sum(PDBdata_dist$Total) * 100, 2)
```




#Working with biomolecular data in R

```{r}
library(bio3d)
pdb <- read.pdb("1hsg")
pdb
```

```{r}
pdb$atom
```


```{r}
pdb$atom[1,"resid"]
```

```{r}
aa321( pdb$atom[pdb$calpha, "resid"] )
```



Q8.  Use the Bio3D write.pdb() function to write out a protein only PDB file for viewing in VMD. Also write out a second separate PDB file for the ligand with residue name MK1

```{r}
protein <- atom.select(pdb, "protein", value = TRUE)
write.pdb(protein, file="1hsg_protein.pdb")
```


Now same for ligand...
```{r}
ligand <- atom.select(pdb, "ligand", value = TRUE)
write.pdb(ligand, file="1hsg_protein.pdb")
```

```{r}
#library(bio3d.view)
#view(ligand)
#view(pdb, "overview", col="sse")
```


```{r}
# Load the package
#pdb <- read.pdb("1hel")
# Normal mode analysis calculation
#modes <- nma(pdb)
#m7 <- mktrj(modes,
 #mode=7,
 #file="mode_7.pdb")

#view(m7, col=vec2color( rmsf(m7) ))
```









