---
title: "Lecture 07 R Functions and Packages"
author: "Reina Bassil"
date: "10/23/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Revisit our functions from last lecutre

```{r}
source("http://tinyurl.com/rescale-R")
```


```{r}
rescale(1:10)
```


```{r}

#Inputting this prints automatic error "Error in x - rng[1] : non-numeric argument to binary operator"
#This isn't very helpful, so we can add a warning/stop function to our function to let users know when things could go/are going wrong.
```



```{r}

```

#Write a function both.na()

##We want to write a function, called both_na(),that counts how many positions in two inputvectors, x and y, both have a missing value

```{r}
# from StackOverflow
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)

is.na(x)
is.na(y)
```


```{r}
which( is.na(x))
```


#Working snippet of code
```{r}
is.na(x) & is.na(y)
which(is.na(x) & is.na(y))
```

#Final function

```{r}
both_na <- function(x,y) {
  sum(is.na(x) & is.na(y))
}
```


```{r}
both_na(x,y)
```

```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
y3 <- c(1, NA, NA, NA, NA, NA)

both_na(x, y1)

both_na(x, y2)

both_na(x, y3)
#this produces no error because y3 has 6 values, and that is a multiple of x which has 3 values
```


```{r}
both_na2 <- function(x, y) {
  if (length(x) != length(y)){
    stop("Error: X and Y are not the same length")
}  
  sum(is.na(x) & is.na(y))
}

```


```{r}

```




#Write a function grade() to determine an overall
grade from a vector of student homework
assignment scores dropping the lowest single
alignment score.
```{r}
# student 1
s1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
# student 2
s2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
# student 3 
s3 <- c(100, NA, NA, NA, NA, NA, NA, NA)
```



```{r}
grade <- function(x) {
  x[is.na(x)] <- 0
    
  drop_lowest <- x[-which.min(x)]
  
  mean(drop_lowest)
}

grade(s3)
```






